An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-24 VLDB2026/main.tex
modify attribute distribution

	Among all algorithms, ACORN-\(\gamma\) shows the highest robustness. During index construction, attribute is mainly used for pruning in the lowest layer graph, while the upper-layer hierarchical navigation graphs remain largely unaffected. As a result, changes in attribute distribution have a limited impact on overall performance.

	\textcolor{cyan}{In contrast, FilteredVamana achieves optimal performance under long-tail and uniform distributions, as these enable construction of balanced label subgraphs with minimal inter-label interference, thus improving search accuracy and efficiency. However, its performance declines under power-law and normal distributions: power-law distributions cause excessive concentration of popular labels and fragmentation of niche labels, disrupting search path connectivity, while normal distributions produce ambiguous graph structures due to label frequency similarities, leading to search divergence. 
	Under the power-law distribution, when the NHQ algorithm constructs an index based on fusion distance, points corresponding to high-frequency attribute values make the graph connections extremely dense, while points corresponding to low-frequency attribute values tend to become isolated due to sparse connections. This extreme imbalance in connectivity severely distorts the structure of the graph, ultimately significantly impairing the overall performance of the algorithm.
	The search performance of VBASE exhibits a direct dependence on attribute selectivity. As selectivity decreases, search speed declines. Notably, only uniform distribution maintains sufficient selectivity among the tested cases; the other three distributions consistently show low selectivity levels that significantly degrade the operational efficiency of VBASE.}

>>>>

	{\color{blue}{
		Algorithms that are highly sensitive to attribute distribution tend to rely passively on the statistical characteristics of the attributes in their design. For instance, graph-based indexing methods like FilteredVamana and NHQ are prone to degraded connectivity or distorted structures under power-law distributions, where label frequencies are highly imbalanced. Similarly, VBASE, which employs a post-filtering strategy, shows efficiency that directly correlates with attribute selectivity—the more selective the filter, the more computational resources are wasted on retrieving irrelevant vectors. FAISS-IVF, which searches within the nearest clusters, may also suffer when relevant data points are unevenly distributed across clusters due to attribute skew, potentially excluding qualified results.

		In contrast, algorithms with low sensitivity to attribute distribution incorporate smarter indexing designs or adaptive query strategies to mitigate these challenges. Milvus, for example, leverages a built-in query optimizer that dynamically determines whether to filter by attributes or search vectors first, based on the selectivity of the query, thus ensuring more robust performance. Other approaches enhance robustness by separating attribute and vector indexing into distinct stages: UNG uses a logical graph to navigate attribute relationships, while CAPS employs an attribute frequency tree to adapt to power-law patterns. ACORN takes a different approach by disregarding attribute distribution entirely during index construction—it builds an ultra-dense, oversaturated graph to ensure that the remaining subgraph, even after aggressive attribute filtering, is still likely to be well-connected and searchable.}}

%	Among all algorithms, ACORN-\(\gamma\) shows the highest robustness. During index construction, attribute is mainly used for pruning in the lowest layer graph, while the upper-layer hierarchical navigation graphs remain largely unaffected. As a result, changes in attribute distribution have a limited impact on overall performance.
%	
%	\textcolor{cyan}{In contrast, FilteredVamana achieves optimal performance under long-tail and uniform distributions, as these enable construction of balanced label subgraphs with minimal inter-label interference, thus improving search accuracy and efficiency. However, its performance declines under power-law and normal distributions: power-law distributions cause excessive concentration of popular labels and fragmentation of niche labels, disrupting search path connectivity, while normal distributions produce ambiguous graph structures due to label frequency similarities, leading to search divergence. 
%	Under the power-law distribution, when the NHQ algorithm constructs an index based on fusion distance, points corresponding to high-frequency attribute values make the graph connections extremely dense, while points corresponding to low-frequency attribute values tend to become isolated due to sparse connections. This extreme imbalance in connectivity severely distorts the structure of the graph, ultimately significantly impairing the overall performance of the algorithm.
%	The search performance of VBASE exhibits a direct dependence on attribute selectivity. As selectivity decreases, search speed declines. Notably, only uniform distribution maintains sufficient selectivity among the tested cases; the other three distributions consistently show low selectivity levels that significantly degrade the operational efficiency of VBASE.}
